How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in \textit{Emys marmorata} (Testudinoidea, Emydidae).

### 2014-11-25 doc/draft_turtle.tex
initial results outline
started filling in results section.

Multinomial logistic regression is an extension of logistic regression, where instead of a binary response there are three or more response classes \citep{Venables2002a}. Similar to the odds ratios calculated from the coefficients of a logistic regression, the relative risk of a classification can be determined from the coefficients of the model. Similar to LDA, the optimal number of PCs as predictors was determined by comparing within-sample AUC values across multiple models.

@@

It is common for some out of sample observations to be misclassified. This misclassification may be due to the model not accurately representing shape variance, systematic differences between the training and test sets, or systematic differences between the accurately and inaccurately classified samples. Testing and training sets are determined completely at random within each class and with respect to shape. Results are not effected by changes in testing or training set assignment.

To determine if there were systematic differences between the correctly and incorrectly classified samples, the multivariate centroids of the correct and incorrect groups were compared to what would be expected by random chance. The group labels were permuted and the difference between the new centroids was calculated. This was repeated 1000 times. The number of permutations less than the empirical difference were counted and divided by 1000, giving a \textit{p}-value. Significant results indicate that correctly and incorrectly classified specimens are systematically different.

@@

@@

\begin{figure}[ht]
  \centering
  \includegraphics{figure/plastra}
  \caption{Depiction of general plastral shape of \textit{E. marmorata} and position of the 19 landmark used in this study. Anterior is towards the top of the figure.}
  \label{fig:plastra}
\end{figure}

>>>>

Multinomial logistic regression is an extension of logistic regression, where instead of a binary response there are three or more response classes \citep{Venables2002a}. Similar to the odds ratios calculated from the coefficients of a logistic regression, the relative risk of a classification can be determined from the coefficients of the model. Similar to LDA, the optimal number of PCs as predictors was determined by comparing within-sample AUC values across multiple models.  % is this true?

@@

It is common for some out of sample observations to be misclassified. This misclassification may be due to the model not accurately representing shape variance, systematic differences between the training and test sets, or systematic differences between the accurately and inaccurately classified samples. Testing and training sets are determined completely at random within each class and with respect to shape. Results were not effected by changes in testing or training set assignment.

To determine if there were systematic differences between the correctly and incorrectly classified samples, the multivariate centroids of the correct and incorrect groups were compared to what would be expected by random chance. The group labels were permuted and the difference between the new centroids was calculated. This was repeated 1000 times. The number of permutations less than the empirical difference were counted and divided by 1000, giving a \textit{p}-value. Significant results indicate that correctly and incorrectly classified specimens are systematically different. This was done only when there were 10 or of that

@@

\subsection{Unsupervised learning}
Comparison of gap statistic values from PAM clustering show that the best number of clusters is 1. This means that there is no means of naturally partitioning plastron shape into distinct subgroups.
% PAM stuff
% figure

\subsection{Supervised learning}
% model selection/fit
For multinomial logistic regression, the best number of predictors was always the first four PCs. This was not the case for either of the LDA or random forest approaches which varied from two to six different PCs. Note that the best random forest models were determined via recursive feature selection, so PCs were not included in order of percent variance explained.

% tabel of RFS choice for random forest

% generalization
AUC values from model generalization, or estimating testing data set membership, does indicate a clear ``best'' classification scheme (Fig. \ref{fig:gen_hist}). While the scheme with two species has the greatest AUC point estimate for each modeling approach, this scheme is not significantly greater from any other except in some limited cases (e.g. LDA).

% misfits
% need to make a table/graph to demonstrate this
Differences in mean shape between correctly and incorrectly classified observations from test set were frequently statistically significant, though there are exceptions. The frequency of these results, however, is important because this means that the different models are poor predictors of class membership. This may be because differences in plastron shape do not align with the any of the hypothesized classification schemes.

\section{Discussion}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/plastra}
  \caption{Depiction of general plastral shape of \textit{E. marmorata} and position of the 19 landmark used in this study. Anterior is towards the top of the figure.}
  \label{fig:plastra}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/gen_res}
  \caption{Bootstrap distributions for genearlized AUC values for each of the classification schemes. Each row corresponds to a different modeling approach: LDA, LDA using best variables from random forest, multinomial logistic regression, and random forest. Each distribution corresponds to 1000 bootstrap replicates.}
  \label{fig:gen_hist}
\end{figure}
